Developing a preprint culture in biology

### 2012-10-08 ms.tex
editing

The first and most often discussed advantage of arXiv and open preprints is
speed (Figure~\ref{fig:map}). The time between submission and the official
publication of a manuscript can be measured in months, sometime in years. For
all this time, the research is known only to a select few: colleagues, editors,
reviewers. Thus, the science cannot be used, discussed, or reviewed by the wider
scientific community. In a recent blog post, C. Titus Brown noted how posting a
paper on arXiv quickly led to a citation (arXiv papers can be cited) and his
research was used by another researcher \cite{bro12}. The current system of
hiding manuscripts before acceptance pose problems for both scientists and
publishers. Manuscripts that are unknown cannot be used and thus take more time
to be cited. It has been shown that high-energy physics, with its high arXiv
submission rate, had the highest immediacy among physics and mathematics
\cite{pra05}. Immediacy measures how quickly articles are cited.

Posting manuscripts as preprints also has the potential to result in higher
quality science, by allowing for pre-publication feedback from a large pool of
reviewers. Prepublication reviews by a small network of colleagues are common in
the biological sciences and are an important part of the scientific process.
These "friendly" reviews increases the chances of errors being caught prior to
publication, allow errors to be identified prior to spending months in review,
and allows authors to improve how they communicate the results of their research
by modifying their writing in response to early feedback. Preprint servers offer
a way to extend this network of colleagues to the entire scientific community,
increasing the number of opportunities for review and revision prior to
publication, and resulting in higher quality submissions and publications.  As
such, preprints can be seen as an integral aspect of a vigorous peer-review
process \cite{hoc12}.

Furthermore, the formal peer-review process as a whole is critically
over-loaded. As the number of active scientists increases and the pressure to
publish increases, it has become increasingly difficult for journals to find
reviewers \cite{hoc09}.  At the same time, rejection rates are high in most
journals \cite{aar08,roh09}, and when not invited to submit a revision, authors
must start the process over again at another journal. As a result, initiatives
to reduce time from submission to publication have emerged across the scientific
community. Rohr et al.  \cite{roh09} called for the recycling and reuse of
peer-reviews: by attaching previous reviews and detailed replies to a new
submission, both the editor and the referees can gauge the work done on the
manuscript, and perhaps evaluate it with less prejudice. In a similar way, the
\emph{Peerage of Science} initiative allows authors to seek anonymous pre-review
by their peers. Some journals now accept to publish papers which received good
evaluations, with \emph{Animal Biology} having recently accepted first a paper
reviewed entirely with the \emph{Peerage of Science} \cite{abb12}, effectively
outsourcing the review process. A widespread use of preprint servers can achieve
the same goal of reducing the time spent in review. By allowing open comments
and criticisms prior to submission, the authors will receive valuable feedback
and can improve the version which will be submitted. With a rich enough
community of scientists depositing preprints, and commenting on them, the
process of an open pre-review can become widespread and will overall increase
the quality of first submissions.

@@

Finally, public preprint servers offer a fair way to establish intellectual
priority by making the work available when immediately after it is complete.
Some manuscripts will spend much more time in the review process and/or in
production after acceptance, than others. This means that
that publication and acceptance dates do not accurately characterize who
came up with an idea first. For this reason, mathematicians and physicists
have embraced arXiv in part to establish priority in a fair way \cite{gin11,cal12}.

@@

In contrast to other disciplines, the field of biology has effectively no preprint
culture, with the exception of small pockets of primarily highly quantitative
research (e.g., epidemiology). While, submitting to preprint servers has become
more common in the past few years, and the quantitative biology section in arXiv
is experiencing rapid growth in submissions \cite{cal12}, the number of biology
papers submitted to preprint servers still represents only a small fraction of
the total research produced in biology.

There are a number of reasons that biologists have not traditionally had a
culture of sharing preprints, many of which are based on common misconceptions
regarding the costs and benefits of sharing preprints. For example, in contrast
to other fields there is a perception in biology that public preprints
make it easier to steal ideas. However, there is no evidence of this happening
in the numerous other fields that have adopted preprint servers, and since preprint
servers create a clear record of who had the idea first, and when, this
appears to be a largely unfounded concern. In other fields preprints serve the
opposite role, they allow straightforward establishment of precendence, letting
researchs lay claim to an idea thus prevening it from being "stolen".

Another major concern is that if a researcher posts a public preprint that they
will not be allowed to submit their paper to their journal of choice. This is
based on a certain interpretation of the Ingelfinger rule: scientists should not
publish the same manuscript twice \cite{alt96}. However, preprints are not
peer-reviewed papers and posting to a public server does not constitute
publication in a peer-reviewed journal.  A preprint is simply a document that
allows ideas to spread and be discussed, but it is not yet formally validated by
the peer-review system. This is why the majority of publishers do not see arXiv
and similar services as a violation of the Ingelfinger rule. Almost all of the
major publishers in biology are preprint-friendly, including: Nature Publishing
Group, PLOS, BMC, PNAS, Science (mostly) \ref{table:policies}, and all the
journals from Elsevier and Springer.  The Ecological Society of America and the
Genetics Society of America also recently changed their policies to allow public
preprints.  Few scientific publications will not consider a manuscript submitted
to a public preprint server. In fact, \emph{Nature} responded to the rumour that
they refused manuscripts submitted to arXiv by saying that ``\emph{Nature} never
wishes to stand in the way of communication between researchers. We seek rather
to add value for authors and the community at large in our peer review,
selection and editing'' \cite{nat05}. Still, a few journals adopt a ``by
default'' hostile attitude towards preprints, mostly due to the lack of clear
policy of the publishers, or perhaps because a preprint culture has not
developed in biology and the practice is still considered unusual. As an
example, Wiley-Blackwell, which publishes some of the leading journals in
biology, has no official policy on the matter.

@@

PeerJ (\href{https://peerj.com/}{https://peerj.com/}) is a new Open Access
publisher that combines both a preprint server, and a peer reviewed journal.
It is focused on the biological and medical sciences, which may help overcome the
perception that preprints do not have a place in biology. Like figshare this is
an advantage relative to arXiv for biologists doing non-quantitative work.
Also like figshare PeerJ allows commenting on posted preprints, improving the
potential for pre-publication
dialog. In addition, preprints can be made private if the authors choose, and
shared only with selected colleagues. While this reduces many of the benefits of
preprints described above, it may allow some researchers who would not otherwise
post preprints to begin to explore the possibility in a manner appropriate to
their current circumstances.

@@

F1000Research is not a public preprint server like the previous three servers.
Whereas arXiv, figshare, and PeerJ offer an option to submit a manuscript
without having it reviewed, papers submitted to F1000Research will eventually be
reviewed. Thus, F1000Research offers a hybrid model with publicly available
manuscripts at time of submission and standard peer-reviews that occur as part
of the submission process. Manuscripts are
considered ``accepted'' and will only be indexed after two positive referee
responses. F1000Research works closely with data providers to integrate raw
data to the paper. For instance, upon submitting a paper, authors are asked to
upload their data, which are then integrated in \emph{e.g.} figshare widgets,
the DOI of which are given in the paper when the data are first mentionned.
By integrating data to the paper, F1000Research is working to make
science more reproducible and open.

@@

The ongoing discussions on the publication process, peer-reviewing and
alternative publication models are all symptoms of the current uneasiness of
scientists with the ever growing obsession with bibliographic metrics such as
the impact factor \cite{Fisher2012}. There is pressure on researchers, wether
they are junior looking for an academic position or senior trying to secure
funding for their team, to orient their publication strategy in order to
maximize their number of publications and total citations. A well-known
consequence is to submit manuscripts first to the most prestigious journals, and
then resubmit to lower level journals as they are rejected. The numerous
negative impacts of such behavior have been discussed in depth \cite{hoc09} and
include a long delay between the time a manuscript is finished to its
publication.  They all contribute somehow in a general slowing down of
scientific progress.  Research activities and the publication process are
drifting away from their fundamental object, namely the diffusion of novel
scientific discoveries.

Developing a preprint culture in biology is not a final solution to the impact
factor inflation, it might however reduce significantly its negative
consequences. Preprints speed up publication, openly, with no judgement on
pertinence and originality. Paradoxically, because peer-reviewing is conducted
following this step in the process of a scientific communication, it might bring
it closer to its fundamental objective. The role of peer-reviewing is to judge
the scientific quality of a study, the match between its hypotheses, methods,
results and conclusions. The peer review process is the first barrier against
fraudulent or bad quality science, susceptible to impede scientific progress. It
is however now dominated by subjective evaluations, where trendy contributions
are promoted to increase the immediacy impact of the journal. Most journals have
in their instructions for reviewers a section devoted to the evaluation of
suitability for the journal, a task usually incumbent to editors. Preprints are
by definition not evaluated, neither for their quality nor their pertinence.
Technically, the difference between a preprint and a traditional publication
should be that the latter as the approval stamp by pairs on its quality. The
relevance of a study, its contribution to science, should only be judged
post-publication by many more readers than the typical two-four anonymous
reviewers. With a such a shift in the diffusion strategy, the role of
traditional journals and their editors would be to showcase scientific
discoveries for specialized readership. Such a process should improve even
further their diffusion, not impede it.

The current opening of the publication process we are currently witnessing with
the arrival of new preprint systems and innovative publication models such as
F1000, PLOS One, Nature Scientific Reports, might significantly bypass the
inflation of bibliographic metrics. They not only promote scientific communications
by changing the focus of peer-reviewing, they also favor the publication of
results that might otherwise be rejected because of low interest, negative
results, descriptive and quantitative studies and datasets. One of drawback of
making publication easier is obviously the proliferation of studies of uneven
quality. A tradeoff between the intensity of the peer-review filtering and the
benefits to science has been hypothesized \cite{Aarssen2012}. With increasingly
stringent peer reviewing, the quality of published papers and easiness of
finding discoveries might increase, at the cost of censorship, increasing load
on authors and reviewers and time for publication. Preprints are simply
bypassing this model, for what we believe is the progress of science: they speed
up the dissemination of scientific discoveries, impede censorship and put on
reader's shoulders the responsibility to judge originality and pertinence.

>>>>

The first and most often discussed advantage of arXiv and open preprints is
speed (Figure~\ref{fig:map}). The time between submission and the official
publication of a manuscript can be measured in months, sometimes in years. For
all this time, the research is known only to a select few: colleagues, editors,
reviewers. Thus, the science cannot be used, discussed, or reviewed by the wider
scientific community. In a recent blog post, C. Titus Brown noted how posting a
paper on arXiv quickly led to a citation (arXiv papers can be cited) and his
research was used by another researcher \cite{bro12}. The current system of
hiding manuscripts before acceptance pose problems for both scientists and
publishers. Manuscripts that are unknown cannot be used and thus take more time
to be cited. It has been shown that high-energy physics, with its high arXiv
submission rate, had the highest immediacy among physics and mathematics
\cite{pra05}. Immediacy measures how quickly articles are cited.

Posting manuscripts as preprints also has the potential to result in higher
quality science, by allowing for pre-publication feedback from a large pool of
reviewers. Prepublication reviews by a small network of colleagues are common in
the biological sciences and are an important part of the scientific process.
These "friendly" reviews increases the chances of errors being caught prior to
publication. Furthermore, the formal peer-review process as a whole is
critically over-loaded. As the number of active scientists increases and the
pressure to publish increases, it has become increasingly difficult for journals
to find reviewers \cite{hoc09}.  At the same time, rejection rates are high in
most journals \cite{aar08,roh09}, and when not invited to submit a revision,
authors must start the process over again at another journal. As a result,
initiatives to reduce time from submission to publication have emerged across
the scientific community. Rohr et al.  \cite{roh09} called for the recycling and
reuse of peer-reviews: by attaching previous reviews and detailed replies to a
new submission, both the editor and the referees can gauge the work done on the
manuscript, and perhaps evaluate it with less prejudice. A widespread use of
preprint servers can achieve the same goal of reducing the time spent in review.
With a rich enough community of scientists depositing preprints, and commenting
on them, the process of an open pre-review can become widespread and will
overall increase the quality of first submissions \cite{hoc12}.

@@

Finally, public preprint servers offer a fair way to establish intellectual
priority by making the work available when it is complete.  Some manuscripts
will spend much more time in the review process and/or in production after
acceptance, than others. This means that publication and acceptance dates do not
accurately characterize who came up with an idea first.  For this reason,
mathematicians and physicists have embraced arXiv in part to establish priority
in a fair way \cite{gin11,cal12}.

@@

In contrast to other disciplines, the field of biology has effectively no
preprint culture, with the exception of small pockets of primarily highly
quantitative research (e.g., epidemiology). Submitting to preprint servers has
become more common in the past few years, and the quantitative biology section
in arXiv is experiencing the fastest growth in submissions \cite{cal12}. Yet,
the number of biology papers submitted to preprint servers still represents only
a small fraction of the total research produced in biology.

There are a number of reasons that biologists have not traditionally had a
culture of sharing preprints, many of which are based on common misconceptions
regarding the costs and benefits of sharing preprints. For example, in contrast
to other fields there is a perception in biology that public preprints
make it easier to steal ideas. However, there is no evidence of this happening
in the numerous other fields that have adopted preprint servers, and since preprint
servers create a clear record of who had the idea first, and when, this
appears to be a largely unfounded concern. In other fields preprints serve the
opposite role, they allow straightforward establishment of precedence, letting
research lay claim to an idea thus preventing it from being "stolen".

Another major concern is that if a researcher posts a public preprint that they
will not be allowed to submit their paper to their journal of choice. This is
based on a certain interpretation of the Ingelfinger rule: scientists should not
publish the same manuscript twice \cite{alt96}. A preprint is simply a document
that allows ideas to spread and be discussed, it is not yet formally validated
by the peer-review system. This is why the majority of publishers do not see
arXiv and similar services as a violation of the Ingelfinger rule. Almost all of
the major publishers in biology are preprint-friendly, including: Nature
Publishing Group, PLOS, BMC, PNAS, Elsevier, and Springer \ref{table:policies}.
The Ecological Society of America and the Genetics Society of America also
recently changed their policies to allow public preprints. In fact,
\emph{Nature} responded to the rumour that they refused manuscripts submitted to
arXiv by saying that ``\emph{Nature} never wishes to stand in the way of
communication between researchers. We seek rather to add value for authors and
the community at large in our peer review, selection and editing'' \cite{nat05}.
Still, a few journals adopt a ``by default'' hostile attitude towards preprints,
mostly due to the lack of clear policy of the publishers, or perhaps because a
preprint culture has not developed in biology and the practice is still
considered unusual. As an example, Wiley-Blackwell, which publishes some of the
leading journals in biology, has no official policy on the matter.

@@

PeerJ (\href{https://peerj.com/}{https://peerj.com/}) is a new Open Access
publisher that combines both a preprint server, and a peer reviewed journal.  It
is focused on the biological and medical sciences, which may help overcome the
perception that preprints do not have a place in biology. Like figshare this is
an advantage relative to arXiv for biologists doing non-quantitative work.  Also
like figshare, PeerJ allows commenting on posted preprints, improving the
potential for pre-publication dialog. In addition, preprints can be made private
if the authors choose, and shared only with selected colleagues. While this
reduces many of the benefits of preprints described above, it may allow some
researchers who would not otherwise post preprints to begin to explore the
possibility in a manner appropriate to their current circumstances.

@@

F1000Research is not a public preprint server like the previous three servers.
Whereas arXiv, figshare, and PeerJ offer an option to submit a manuscript
without having it reviewed, papers submitted to F1000Research will eventually
be reviewed. Thus, F1000Research offers a hybrid model with publicly available
manuscripts at time of submission and standard peer-reviews that occur as part
of the submission process. Manuscripts are considered ``accepted'' and will
only be indexed after two positive referee responses. F1000Research works
closely with data providers to integrate raw data to the paper. For instance,
upon submitting a paper, authors are asked to upload their data, which are then
integrated in \emph{e.g.} figshare widgets, the DOI of which are given in the
paper when the data are first mentioned.  By integrating data to the paper,
F1000Research is working to make science more reproducible and open.

@@

The ongoing discussions on the publication process, peer-reviewing and
alternative publication models are all symptoms of the current uneasiness of
scientists with the ever growing obsession with bibliographic metrics such as
the impact factor \cite{Fisher2012}. There is pressure on researchers to orient
their publication strategy in order to maximize their number of publications and
total citations. A well-known consequence is to submit manuscripts first to the
most prestigious journals, and then resubmit to ``lower level'' journals as they
are rejected. The numerous negative impacts of such behavior have been discussed
in depth \cite{hoc09} and include a long delay between the time a manuscript is
finished to its publication.  They all contribute somehow in a general slowing
down of scientific progress.  Research activities and the publication process
are drifting away from their fundamental object, namely the diffusion of novel
scientific discoveries.

Developing a preprint culture in biology will not solve all problems with the
current publication process. It might however reduce significantly its negative
consequences. Preprint servers speed up publication and make no judgement on
pertinence and originality. Paradoxically, because peer-reviewing is conducted
\emph{after} this step in the process of scientific communication, it might
bring it closer to its fundamental objective. The role of peer-reviewing is to
judge the scientific quality of a study. The peer review process is the first
barrier against the fraudulent and poor quality science susceptible to impede
scientific progress. However, it is now dominated by subjective evaluations,
where trendy contributions are promoted to increase impact factors. Preprints
are by definition not evaluated, neither for their quality nor their pertinence.
Technically, the difference between a preprint and a traditional publication
should be that the latter as the approval stamp by pairs on its quality. The
relevance of a study, its contribution to science, should only be judged
post-publication by many more readers than the typical two-four anonymous
reviewers. With a such a shift in the diffusion strategy, the role of
traditional journals and their editors would be to showcase scientific
discoveries for specialized readership. Such a process should improve even
further their diffusion, not impede it.

One of drawback of making publication easier is the proliferation of studies of
uneven quality. A trade-off between the intensity of the peer-review filtering
and the benefits to science has been hypothesized \cite{Aarssen2012}.  With
increasingly stringent peer reviewing, the quality of published papers and
easiness of finding discoveries might increase, at the cost of censorship, an
increased load on authors and reviewers, and more delays for publication.
Preprints are simply bypassing this model for what we believe is the progress of
science: they speed up the dissemination of scientific discoveries, impede
censorship and put on reader's shoulders the responsibility to judge originality
and pertinence.
